---
title: "Flow Plots"
description: |
  FF Metrics and visualizations
author:
  - name: R. Peek 
date: "Updated `r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(layout="l-page")
library(here)
library(glue)
```

# Pulling Full Natural Flow Data

First we pulled data from CDEC for Station TLG (Tuolumne at La Grange). These data were full natural flow calculations from DWR.

## Data Cleaning

Many data points were negative, which creates downstream issues with the functional flow analysis. Thus, these values were converted to NAs and then imputed using a structural time series model (Kalman model) with the {`imputeTS`} package in R.

We then calculated daily and annual statistics on the full period of record, including centroid timing, annual flow volumes, and the percentiles associated with daily and annual flows.

### Data Summary

Annual flow stats were calculated using the {`fasstr`} package. Unless otherwise specified, flow is currently in **`cms`**.

```{r flowstats-ann, layout="l-page"}


# y axis: thousand acre feet / year
# percentile of metric
suppressPackageStartupMessages(source(here("scripts/functions/f_dwr_03_load_fnf.R")))
f_dwr_03_load_fnf()

library(DT)
# filter to just some basic columns
dwr_ann_df %>% 
  select(station_id:Annual_Maximum, starts_with("DoY"), tot_vol_acft:ann_vol_percentile) %>% 
  # round data
  mutate(across(.cols = Annual_Maximum:ann_vol_percentile, round, 1)) %>% 
  DT::datatable()
```

Daily flow stats are in both cms and cfs, and volumes have been converted to thousand acre-feet. This is only a portion of the total data.

```{r flowstats-daily}
dwr_daily_df_filt <- dwr_daily_df %>% filter(WY>1986 & WY < 2022)

# summarize filtered
summary(dwr_daily_df_filt)

```

## Visualization

We can visualize the data in a variety of ways. The DWR daily data from the TLG site has many negative values.

### Negative value distribution of raw data:

```{r negflow, fig.cap="Raw Full Natural Flow plots"}

library(ggplot2)
library(plotly)

siteID <- "TLG"
# visualize the negatives
g1 <- dwr_daily_df_filt %>% 
  # replace NAs with -1 to plot
  mutate(flow_rev = ifelse(is.na(flow), -1, flow)) %>% {
    ggplot() +
      geom_col(data = ., aes(x=date, y=flow_rev),
               color=ifelse(.$flow_rev >0, "cyan4", "red3")) +
      labs(title=glue("DWR Full Natural Flow for {siteID}: Raw Data"),
           caption="Negative values are in red",
           x="", y="Flow (cfs)") +
      theme_classic() +
      scale_y_continuous(labels = scales::comma)}
ggplotly(g1)

```

To interpolate these data, we converted all negative values to NA, and then imputed data using a structural time series model (Kalman Smoother) with the {`imputeTS`} package in R. This performs well and provides a tighter relationship than using a 7-day moving average (centered). Both methods have been compared, but for calculating the function flow metrics we opted to move forward with the Kalman smoothed interpolation.

### Plot the Interpolated Data

```{r interpDat, fig.cap="DWR FNF for TLG: Negative values are in red, Kalman Smoothed data in yellow"}

g1 <- dwr_daily_df_filt %>% 
  # replace NAs with -1 to plot
  mutate(flow_rev = ifelse(is.na(flow), -1, flow)) %>% {
    ggplot(data=.) +
      geom_col(aes(x=date, y=flow_rev),
               color=ifelse(.$flow_rev >0, "cyan4", "red3")) +
      labs(title=glue("DWR Full Natural Flow for {siteID}"),
           subtitle="Negative values are in red, Kalman Smoothed data in yellow",
           x="", y="Flow (cfs)") +
      geom_point(aes(x=date, y=flow_interp_cfs), fill="yellow", pch=21) +
      theme_classic() +
      scale_y_continuous(labels = scales::comma)}
ggplotly(g1)

```


## Flow Plots

Below we show many different ways to visualize the flow and flow volumes.

### Plot Annual Flow Timing

This shows the center of timing (50%) and the 25%, 33% and 75%. This is the day of the year that a given percentage of the total volume has occurred. Note 1997 was very early as a very wet year, while 2015 through 2018 all clustered together because while there was less total water, that center of timing still happened earlier.

```{r annflow, fig.cap="Flow Timing of Total Annual Volume"}

library(fasstr)
# flow timing   
plot_annual_flow_timing(data = dwr_daily_df, dates = date, 
                        values = flow_interp_cms, groups = station_id,
                        water_year_start = 10, 
                        start_year = 1987, end_year = 2021, 
                        percent_total = c(25, 33.3, 50, 75))

```


```{r dailyflow,  eval =FALSE, echo=FALSE, fig.cap="Daily flow statistics for period of record"}
# Add a low flow year (2014) for comparison.

# flow timing   
plot_daily_stats(data = dwr_daily_df, dates = date, 
                   values = flow_interp_cms, groups = station_id,
                   water_year_start = 10, 
                   start_year = 1987, end_year = 2021, add_year = 2014)

```


```{r cumulativeStats, echo = FALSE, eval=FALSE, fig.cap="Cumulative Flow statistics"}

plot_daily_cumulative_stats(data = dwr_daily_df, dates = date, 
                              values = flow_interp_cms, groups = station_id,
                              water_year_start = 10, 
                              start_year = 1987, end_year = 2021)

```

<!--
> Flow Duration Curves for Period of Record

```{r FlowDurationCurves, fig.cap="Cumulative Flow statistics"}

plot_flow_duration(data = dwr_daily_df, dates = date, 
                              values = flow_interp_cms, groups = station_id,
                              water_year_start = 10, 
                              start_year = 1987, end_year = 2021)

```


### Flow Summary for Period of Record

Here we show the interpolated data vs. the raw data (with negative values). In some cases this is certainly an overestimation, but in order to calculate flow metrics we need a complete (and positive) flow record. This is just one approach.

```{r totalPOR, fig.cap="Flow for Period of Record", eval=FALSE}

siteID <- "TLG"

# quick plot
p1 <- ggplot(data = dwr_daily_df_filt) +
  geom_col(aes(x=date, y=flow),color="cyan4") +
  geom_point(aes(x=date, y=flow_interp_cfs), pch=21, fill="yellow", size=2) +
  labs(
    title=glue("DWR Full Natural Flow for {siteID}: Interpolated {min(dwr_daily_df_filt$WY)}-{max(dwr_daily_df_filt$WY)}"),
    caption="Yellow dots represent negative values that interpolated with Kalman structural time series {imputeTS}",
    x="", y="Flow (cfs)") +
  scale_y_continuous(labels = scales::comma)+
  theme_classic()
ggplotly(p1)

```
-->

## Functional Flow Metrics (FFM) (DWR TLG)

Next we can read in the data from the functional flow metrics and bind with our annual stats (centroid timing, etc). We also calculate what percentiles each metric is across the period of record for that metric (so essentially ranking the data). We joined these data with a tercile water year type calculated stored on the TNC website.

```{r joindata, echo=FALSE}

ffm <- read_csv(here("output/ffc_TLG/2823750_ffc_results.csv"), show_col_types = FALSE) %>% 
  #drop first column and peak metrics
  dplyr::select(-1, -starts_with("Peak_"))

# select ann metrics we want
dwr_ann_df_sel <- select(dwr_ann_df, Year:Annual_Minimum, starts_with("DoY"),tot_vol_acft, ann_vol_percentile)

# join
dat_out <- left_join(ffm, dwr_ann_df_sel)

# function to calc percentiles
get_percentiles <- function(x){
  vecout <- round((stats::ecdf(x))(x), 4) * 100
  return(vecout)
}

# run it on all columns!
dat_out_prct <- dat_out %>% 
  mutate(across(DS_Dur_WS:Wet_Tim, ~get_percentiles(.x), .names= "{.col}_pct"))

```

```{r centTiming, echo=FALSE, eval=FALSE}
# calc own timing:
# to calculate our own timing percentiles, we can use this:
library(fasstr)
ann_flow_timing <- calc_annual_flow_timing(data = dwr_daily_df_filt, dates = date,
                                           values = flow_interp_cms,
                                           groups = station_id,
                                           water_year_start = 10,
                                           start_year = 1987, end_year = 2021,
                                           percent_total = c(25, 50, 75)) %>%
# clean names
    janitor::clean_names() %>%
    rename(dowy_25pct_tot_q = 3, date_25pct_tot_q = 4,
           dowy_50pct_tot_q = 5, date_50pct_tot_q = 6,
           dowy_75pct_tot_q = 7, date_75pct_tot_q = 8)

```

```{r addWYT}

# join with water year types
wyt <- read_rds(here("data_clean/wyt_cdec_tnc_2823750.rds"))

# join with data
dat_out_p <- left_join(dat_out_prct, 
                       wyt %>% select(year, starts_with("wyt")),
                       by=c("Year"="year"))

#levels(dat_out_p$wyt_tnc)
#levels(dat_out_p$wyt_sj_cdec)

```

### Centroid Timing of Total Annual Volume

We can plot the centroid timing at different values (here, 50%, and 75% of the total annual flow volume) over the period of record. There is a general trend for the center of volume to occur earlier than the spring recession timing.

```{r plotcentTiming, fig.cap="Centroid Timing for 50%, 75% vs. Spring Timing"}

# plot
ggplot()+
  geom_point(data=dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ, fill=wyt_sj_cdec), pch=21, size=2.5, col="black") +
  geom_smooth(data=dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ), method = "gam", color="cyan4") +
  geom_point(data=dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ, fill=wyt_sj_cdec), pch=23, size=2.5, col="darkblue")+
  geom_smooth(data=dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ), method = "gam", color="darkblue") +
  geom_abline(intercept = 0, alpha=0.5, lty=2) +
  scale_y_continuous(breaks=c(seq(90,270,20)))+
  theme_bw() +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  #facet_wrap(.~wyt) +
  labs(title = "Centroid Timing vs. Spring Timing (Day of Water Year",
       subtitle = "(darkblue=75%, lightblue=50% of tot vol, 1:1 line is dashed)",
       y="Day of Water Year % Annual Volume Reached",
       x="Spring Timing (Day of water year)")
  

```


```{r plotcentTimingTNC, fig.cap="Centroid Timing for 50%, 75% vs. Spring Timing"}

# plot
(g2 <- ggplot()+
  geom_point(data=dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ, fill=wyt_tnc), pch=21, size=2.5, col="black") +
  geom_smooth(data=dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ), method = "gam", color="cyan4") +
  geom_point(data=dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ, fill=wyt_tnc), pch=23, size=2.5, col="darkblue")+
  geom_smooth(data=dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ), method = "gam", color="darkblue") +
  geom_abline(intercept = 0, alpha=0.5, lty=2) +
  scale_y_continuous(breaks=c(seq(90,270,20)))+
  theme_bw() +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  #facet_wrap(.~wyt) +
  labs(title = "Centroid Timing vs. Spring Timing (Day of Water Year",
       subtitle = "(darkblue=75%, lightblue=50% of tot vol, 1:1 line is dashed)",
       y="Day of Water Year % Annual Volume Reached",
       x="Spring Timing (Day of water year)"))
  

```

### Volume vs. Percentiles: FNF

These plots are the percentiles calculated for functional flow metric for the period of record of station TLG (1987-2021). We calculated the functional flow metrics based on interpolated daily flows, and then took a cumulative distribution function to calculate the percentile value of each calculated FF metric in each year. These can then be plotted against the total annual volume for the same year.


```{r plotVolPercnt-DS, fig.cap="Total Volume vs. Dry Season FFM Percentiles"}

library(cowplot)
library(tidyr)
library(ggthemes)
library(scales)
# get color al
#show_col(colorblind_pal()(8))
colblind8 <-colorblind_pal()(8)

# plot
gg_out <- dat_out_p %>% 
  select(Year, wyt_sj_cdec, wyt_tnc, tot_vol_acft, ann_vol_percentile,
         DS_Dur_WS_pct:Wet_Tim_pct) %>% 
  pivot_longer(DS_Dur_WS_pct:Wet_Tim_pct, 
               names_to = "FFM", 
               values_to = "ffm_por_percentile")
  
gg_out %>% 
  filter(grepl("DS_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[2]) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  #facet_grid(FFM~wyt) + # gams don't work for this
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann. Volume vs. FFM: Dry Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

gg_out %>% 
  filter(grepl("DS_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[2]) +
  #scale_fill_viridis_d("CDEC WYT", option = "C") +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  #facet_grid(FFM~wyt) + # gams don't work for this
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Dry Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")
  
```



```{r plotVolPercnt-FA, fig.cap="Total Volume vs. Fall Pulse FFM Percentiles"}

gg_out %>% 
  filter(grepl("FA_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "glm", color=colblind8[7]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Fall Pulse",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")


gg_out %>% 
  filter(grepl("FA_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "glm", color=colblind8[7]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Fall Pulse",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

```


```{r plotVolPercnt-Wet, fig.cap="Total Volume vs. Wet Season FFM Percentiles"}

gg_out %>% 
  filter(grepl("Wet_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[6]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Wet Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

gg_out %>% 
  filter(grepl("Wet_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[6]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Wet Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")


```


```{r plotVolPercnt-SP, fig.cap="Total Volume vs. Spring Timing FFM Percentiles"}

gg_out %>% 
  filter(grepl("SP_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), 
             pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[4]) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Spring Recession",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

gg_out %>% 
  filter(grepl("SP_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), 
             pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[4]) +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Spring Recession",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

```

## Functional Flow Metrics: Livneh et al.

Next we can read in the data calculated Livneh et al. to generate the same functional flow metrics and annual stats (centroid timing, etc).

```{r getLivneh, echo=FALSE, message=FALSE, warning=FALSE}

# ffm
liv_ffm <- read_csv(here("output/ffc_livneh/2823750_livneh_ffc_results.csv"),
                    show_col_types = FALSE) %>% 
  #drop first column and peak metrics
  dplyr::select(-1, -starts_with("Peak_"))

# get livneh data
source(here("scripts/functions/f_livneh_02_load_fnf.R"))
f_load_livneh_fnf()

# select ann metrics we want
liv_ann_df_sel <- select(liv_ann_df, Year:Annual_Minimum,
                         starts_with("DoY"),tot_vol_acft, ann_vol_percentile)

# join
liv_dat_out <- left_join(liv_ffm, liv_ann_df_sel)

# function to calc percentiles
get_percentiles <- function(x){
  vecout <- round((stats::ecdf(x))(x), 4) * 100
  return(vecout)
}

# run it on all columns!
liv_dat_out_p <- liv_dat_out %>% 
  mutate(across(DS_Dur_WS:Wet_Tim, ~get_percentiles(.x), .names= "{.col}_pct"))

# join with water year types
wyt <- read_rds(here("data_clean/wyt_cdec_tnc_2823750.rds"))

# join with data
liv_dat_out_p <- left_join(liv_dat_out_p, 
                       wyt %>% select(year, starts_with("wyt")),
                       by=c("Year"="year"))

# levels(liv_dat_out_p$wyt_tnc)
# levels(liv_dat_out_p$wyt_sj_cdec)
# range(liv_dat_out_p$Year)

```

### Centroid Timing of Total Annual Volume

We can plot the centroid timing at different values (here, 50%, and 75% of the total annual flow volume) over the period of record. There is a general trend for the center of volume to occur earlier than the spring recession timing.

```{r plotcentTiming2, fig.cap="Centroid Timing for 50%, 75% vs. Spring Timing"}

# plot
ggplot()+
  geom_point(data=liv_dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ, fill=wyt_sj_cdec), pch=21, size=2.5, col="black") +
  geom_smooth(data=liv_dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ), method = "gam", color="cyan4") +
  geom_point(data=liv_dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ, fill=wyt_sj_cdec), pch=23, size=2.5, col="darkblue")+
  geom_smooth(data=liv_dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ), method = "gam", color="darkblue") +
  geom_abline(intercept = 0, alpha=0.5, lty=2) +
  scale_y_continuous(breaks=c(seq(90,270,20)))+
  theme_bw() +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  #facet_wrap(.~wyt) +
  labs(title = "Centroid Timing vs. Spring Timing (Day of Water Year",
       subtitle = "(darkblue=75%, lightblue=50% of tot vol, 1:1 line is dashed)",
       y="Day of Water Year % Annual Volume Reached",
       x="Spring Timing (Day of water year)")
  

```


### Volume vs. Percentiles: FNF

These plots are the percentiles calculated for functional flow metric for the period of record of station TLG (1987-2021). We calculated the functional flow metrics based on interpolated daily flows, and then took a cumulative distribution function to calculate the percentile value of each calculated FF metric in each year. These can then be plotted against the total annual volume for the same year.


```{r plotVolPercnt-DS2, fig.cap="Total Volume vs. Dry Season FFM Percentiles"}

library(cowplot)
library(tidyr)
library(ggthemes)
library(scales)
# get color al
#show_col(colorblind_pal()(8))
colblind8 <-colorblind_pal()(8)

# plot
gg_out_liv <- liv_dat_out_p %>% 
  select(Year, wyt_sj_cdec, wyt_tnc, tot_vol_acft, ann_vol_percentile,
         DS_Dur_WS_pct:Wet_Tim_pct) %>% 
  pivot_longer(DS_Dur_WS_pct:Wet_Tim_pct, 
               names_to = "FFM", 
               values_to = "ffm_por_percentile")
  
gg_out_liv %>% 
  filter(grepl("DS_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[2]) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  #facet_grid(FFM~wyt) + # gams don't work for this
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Dry Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

gg_out_liv %>% 
  filter(grepl("DS_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[2]) +
  #scale_fill_viridis_d("CDEC WYT", option = "C") +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  #facet_grid(FFM~wyt) + # gams don't work for this
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Dry Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")
  
```



```{r plotVolPercnt-FA2, fig.cap="Total Volume vs. Fall Pulse FFM Percentiles"}

gg_out_liv %>% 
  filter(grepl("FA_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "glm", color=colblind8[7]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Fall Pulse",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")


gg_out_liv %>% 
  filter(grepl("FA_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "glm", color=colblind8[7]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Fall Pulse",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

```


```{r plotVolPercnt-Wet2, fig.cap="Total Volume vs. Wet Season FFM Percentiles"}

gg_out_liv %>% 
  filter(grepl("Wet_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[6]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Wet Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

gg_out_liv %>% 
  filter(grepl("Wet_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[6]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Wet Season",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")


```


```{r plotVolPercnt-SP2, fig.cap="Total Volume vs. Spring Timing FFM Percentiles"}

gg_out_liv %>% 
  filter(grepl("SP_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), 
             pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[4]) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Spring Recession",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

gg_out_liv %>% 
  filter(grepl("SP_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_tnc), 
             pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[4]) +
  scale_fill_viridis_d("WYT TNC", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Spring Recession",
       y="Total Volume (thousand acre feet)", 
       x="Percentile of metric")

```