---
title: "Flow Plots"
description: |
  FF Metrics and visualizations
author:
  - name: R. Peek 
date: "Updated `r Sys.Date()`"
output: distill::distill_article
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(layout="l-page")
library(here)
library(glue)
```

# Pulling Full Natural Flow Data

First we pulled data from CDEC for Station TLG (Tuolumne at La Grange). These data were full natural flow calculations from DWR.

## Data Cleaning

Many data points were negative, which creates downstream issues with the functional flow analysis. Thus, these values were converted to NAs and then imputed using a structural time series model (Kalman model) with the {`imputeTS`} package in R.

We then calculated daily and annual statistics on the full period of record, including centroid timing, annual flow volumes, and the percentiles associated with daily and annual flows.

### Data Summary

Annual flow stats were calculated using the {`fasstr`} package. Unless otherwise specified, flow volumes are currently in **`cms`**.

```{r flowstats-ann, layout="l-page"}

suppressPackageStartupMessages(source(here("scripts/functions/f_load_dwr_fnf.R")))
f_load_dwr_fnf()

library(DT)
# filter to just some basic columns
dwr_ann_df %>% 
  select(station_id:Annual_Maximum, starts_with("DoY"), tot_vol_acft:ann_vol_percentile) %>% 
  # round data
  mutate(across(.cols = Annual_Maximum:ann_vol_percentile, round, 1)) %>% 
  DT::datatable()
```

Daily flow stats are in both cms and cfs, and volumes have been converted to acre-feet^3^. This is only a portion of the total data.

```{r flowstats-daily}
dwr_daily_df_filt <- dwr_daily_df %>% filter(WY>1986 & WY < 2022)

# summarize filtered
summary(dwr_daily_df_filt)

```

## Visualization

We can visualize the data in a variety of ways.

### Negative value distribution of raw data:

```{r negflow, fig.cap="Raw Full Natural Flow plots"}

library(ggplot2)
library(plotly)

siteID <- "TLG"
# visualize the negatives
g1 <- dwr_daily_df_filt %>% 
  # replace NAs with -1 to plot
  mutate(flow_rev = ifelse(is.na(flow), -1, flow)) %>% {
    ggplot() +
      geom_col(data = ., aes(x=date, y=flow_rev),
               color=ifelse(.$flow_rev >0, "cyan4", "red3")) +
      labs(title=glue("DWR Full Natural Flow for {siteID}: Raw Data"),
           caption="Negative values are in red",
           x="", y="Flow (cfs)") +
      theme_classic() +
      scale_y_continuous(labels = scales::comma)}
ggplotly(g1)

```

### Plot the NAs in the Data

```{r nasflow, fig.cap="Negative values replaced with NAs to visualize data"}

# visualize the nas and interpolated versions
imputeTS::ggplot_na_distribution(dwr_daily_df_filt$flow_na)

```

### Plot the Interpolated Data

```{r}

g1 <- dwr_daily_df_filt %>% 
  # replace NAs with -1 to plot
  mutate(flow_rev = ifelse(is.na(flow), -1, flow)) %>% {
    ggplot(data=.) +
      geom_col(aes(x=date, y=flow_rev),
               color=ifelse(.$flow_rev >0, "cyan4", "red3")) +
      labs(title=glue("DWR Full Natural Flow for {siteID}: Raw Data"),
           caption="Negative values are in red",
           x="", y="Flow (cfs)") +
      geom_point(aes(x=date, y=flow_interp_cfs), fill="yellow", pch=21) +
      theme_classic() +
      scale_y_continuous(labels = scales::comma)}
ggplotly(g1)

```


## Flow Plots

Below we show many different ways to visualize the flow and flow volumes.

### Plot Annual Flow Timing

```{r annflow, fig.cap="Flow Timing of Total Annual Volume"}

library(fasstr)
# flow timing   
plot_annual_flow_timing(data = dwr_daily_df, dates = date, 
                        values = flow_interp_cms, groups = station_id,
                        water_year_start = 10, 
                        start_year = 1987, end_year = 2021, 
                        percent_total = c(25, 33.3, 50, 75))

```


### Plot Daily Flow Statistics

Add a low flow year (2014) for comparison.

```{r dailyflow, fig.cap="Daily flow statistics for period of record"}

# flow timing   
plot_daily_stats(data = dwr_daily_df, dates = date, 
                   values = flow_interp_cms, groups = station_id,
                   water_year_start = 10, 
                   start_year = 1987, end_year = 2021, add_year = 2014)

```


### Cumulative Flow Stats

```{r cumulativeStats, fig.cap="Cumulative Flow statistics"}

plot_daily_cumulative_stats(data = dwr_daily_df, dates = date, 
                              values = flow_interp_cms, groups = station_id,
                              water_year_start = 10, 
                              start_year = 1987, end_year = 2021)

```

<!--
> Flow Duration Curves for Period of Record

```{r FlowDurationCurves, fig.cap="Cumulative Flow statistics"}

plot_flow_duration(data = dwr_daily_df, dates = date, 
                              values = flow_interp_cms, groups = station_id,
                              water_year_start = 10, 
                              start_year = 1987, end_year = 2021)

```

-->
### Flow Summary for Period of Record

Here we show the interpolated data vs. the raw data (with negative values). In some cases this is certainly an overestimation, but in order to calculate flow metrics we need a complete (and positive) flow record. This is just one approach.

```{r totalPOR, fig.cap="Flow for Period of Record", eval=FALSE}

siteID <- "TLG"

# quick plot
p1 <- ggplot(data = dwr_daily_df_filt) +
  geom_col(aes(x=date, y=flow),color="cyan4") +
  geom_point(aes(x=date, y=flow_interp_cfs), pch=21, fill="yellow", size=2) +
  labs(
    title=glue("DWR Full Natural Flow for {siteID}: Interpolated {min(dwr_daily_df_filt$WY)}-{max(dwr_daily_df_filt$WY)}"),
    caption="Yellow dots represent negative values that interpolated with Kalman structural time series {imputeTS}",
    x="", y="Flow (cfs)") +
  scale_y_continuous(labels = scales::comma)+
  theme_classic()
ggplotly(p1)

```

## Functional Flow Metrics (FFM)

Next we can read in the data from the functional flow metrics and bind with our annual stats (centroid timing, etc). We also calculate what percentiles each metric is across the period of record for that metric (so essentially ranking the data). We joined these data with a tercile water year type calculated stored on the TNC website.

```{r joindata, echo=FALSE}

ffm <- read_csv(here("output/ffc_TLG/2823750_ffc_results.csv"), show_col_types = FALSE) %>% 
  #drop first column and peak metrics
  dplyr::select(-1, -starts_with("Peak_"))

# select ann metrics we want
dwr_ann_df_sel <- select(dwr_ann_df, Year:Annual_Minimum, starts_with("DoY"),tot_vol_acft, ann_vol_percentile)

# join
dat_out <- left_join(ffm, dwr_ann_df_sel)

# function to calc percentiles
get_percentiles <- function(x){
  vecout <- round((stats::ecdf(x))(x), 4) * 100
  return(vecout)
}

# run it on all columns!
dat_out_p <- dat_out %>% 
  mutate(across(DS_Dur_WS:Wet_Tim, ~get_percentiles(.x), .names= "{.col}_pct"))

## to calculate our own timing percentiles, we can use this:
# library(fasstr)
# ann_flow_timing <- calc_annual_flow_timing(data = dwr_daily_df_filt, dates = date, 
#                                            values = flow_interp_cms, 
#                                            groups = station_id,
#                                            water_year_start = 10, 
#                                            start_year = 1987, end_year = 2021, 
#                                            percent_total = c(25, 50, 75)) %>%
# # clean names
#     janitor::clean_names() %>% 
#     rename(dowy_25pct_tot_q = 3, date_25pct_tot_q = 4,
#            dowy_50pct_tot_q = 5, date_50pct_tot_q = 6,
#            dowy_75pct_tot_q = 7, date_75pct_tot_q = 8)


# TNC join with water year types:
wyt <- read_csv(here("data_clean/tnc_wyt_2823750.csv"), show_col_types = FALSE)

dat_out_p <- left_join(dat_out_p, wyt[,c(1,2,5)], by=c("Year"="year")) %>% 
  rename(wyt_tnc = wyt)

# get CDEC wyt
cdec_wyt <- wateRshedTools::ca_wytypes %>% 
  select(WY, sj_WYtype)

dat_out_p <- left_join(dat_out_p, cdec_wyt, by=c("Year"="WY")) %>%
  rename(wyt_sj_cdec = sj_WYtype) %>% 
  # fill one NA with "D"
  mutate(wyt_sj_cdec = ifelse(is.na(wyt_sj_cdec), "D", wyt_sj_cdec)) %>% 
  mutate(wyt_sj_cdec = 
           forcats::fct_relevel(wyt_sj_cdec, c("W","AN","BN", "D","C")))

# levels(dat_out_p$wyt_sj_cdec)
```

### Centroid Timing of Total Annual Volume

We can plot the centroid timing at different values (here, 33%, 50%, and 75% of the total annual flow volume) over the period of record. There is a general trend for the center of volume to occur earlier than the spring recession timing.

```{r plotcentTiming, fig.cap="Centroid Timing for 33%, 50%, 75% vs. Spring Timing"}

# plot
ggplot()+
  geom_point(data=dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ, fill=wyt_sj_cdec), pch=21, size=2.5, col="black") +
  geom_smooth(data=dat_out_p, aes(x=SP_Tim, y=DoY_50pct_TotalQ), method = "gam", color="cyan4") +
  geom_point(data=dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ, fill=wyt_sj_cdec), pch=23, size=2.5, col="darkblue")+
  geom_smooth(data=dat_out_p, aes(x=SP_Tim, y=DoY_75pct_TotalQ), method = "gam", color="darkblue") +
  geom_abline(intercept = 0, alpha=0.5, lty=2) +
  scale_y_continuous(breaks=c(seq(90,270,20)))+
  theme_bw() +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  #facet_wrap(.~wyt) +
  labs(title = "Centroid Timing vs. Spring Timing (Day of Water Year",
       subtitle = "(darkblue=75%, lightblue=50% of tot vol, 1:1 line is dashed)",
       y="Day of Water Year % Annual Volume Reached",
       x="Spring Timing (Day of water year)")
  

```


### Volume vs. Percentiles


```{r plotVolPercnt-DS, fig.cap="Total Volume vs. Dry Season FFM Percentiles"}

library(cowplot)
library(tidyr)
library(ggthemes)
library(scales)
# get color al
#show_col(colorblind_pal()(8))
colblind8 <-colorblind_pal()(8)

# plot
gg_out <- dat_out_p %>% 
  select(Year, wyt_sj_cdec, tot_vol_acft, ann_vol_percentile,
         DS_Dur_WS_pct:Wet_Tim_pct) %>% 
  pivot_longer(DS_Dur_WS_pct:Wet_Tim_pct, 
               names_to = "FFM", 
               values_to = "ffm_por_percentile")
  
gg_out %>% 
  filter(grepl("DS_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[2]) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  #facet_grid(FFM~wyt) + # gams don't work for this
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Dry Season",
       y=bquote("Total Volume" ~('acre ft'^3)), 
       x="Percentile of each year")
  
```



```{r plotVolPercnt-FA, fig.cap="Total Volume vs. Fall Pulse FFM Percentiles"}

gg_out %>% 
  filter(grepl("FA_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "glm", color=colblind8[7]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Fall Pulse",
       y=bquote("Total Volume" ~('acre ft'^3)), 
       x="Percentile of each year")

```


```{r plotVolPercnt-Wet, fig.cap="Total Volume vs. Wet Season FFM Percentiles"}

gg_out %>% 
  filter(grepl("Wet_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[6]) +
  facet_wrap(.~FFM) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Wet Season",
       y=bquote("Total Volume" ~('acre ft'^3)), 
       x="Percentile of each year")

```


```{r plotVolPercnt-SP, fig.cap="Total Volume vs. Spring Timing FFM Percentiles"}

gg_out %>% 
  filter(grepl("SP_", FFM)) %>% 
  ggplot(data=.) +
  geom_point(aes(x=ffm_por_percentile, y=tot_vol_acft, fill=wyt_sj_cdec), 
             pch=21, size=2.5) +
  geom_smooth(aes(x=ffm_por_percentile, y=tot_vol_acft), method = "gam", color=colblind8[4]) +
  scale_fill_viridis_d("CDEC WYT", option = "C") +
  facet_wrap(.~FFM) +
  theme_cowplot(font_size = 12) +
  cowplot::background_grid(major = "xy") +
  labs(subtitle = "Total Ann Volume vs. FFM: Spring Recession",
       y=bquote("Total Volume" ~('acre ft'^3)), 
       x="Percentile of each year")

```

